---
title: "RandomForest"
author: "Diego Polo"
date: "4 de enero de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TITANIC SURVIVAL FORECASTING USING RANDOMFOREST ALGORITHM

This is my first kernel and I decided to use RamdonForest algorithm as many of features of datasets are factors and I think RandomForest clasificator would work fine

Overview

The data has been split into two groups:

* training set (train.csv)
* test set (test.csv)

The training set should be used to build your machine learning models. For the training set, we provide the outcome (also known as the “ground truth”) for each passenger. Your model will be based on “features” like passengers’ gender and class. You can also use feature engineering to create new features.

The test set should be used to see how well your model performs on unseen data. For the test set, we do not provide the ground truth for each passenger. It is your job to predict these outcomes. For each passenger in the test set, use the model you trained to predict whether or not they survived the sinking of the Titanic.

We also include gender_submission.csv, a set of predictions that assume all and only female passengers survive, as an example of what a submission file should look like.

###Data Dictionary
VariableDefinitionKey survival Survival 0 = No, 1 = Yes pclass Ticket class 1 = 1st, 2 = 2nd, 3 = 3rd sex Sex Age Age in years sibsp # of siblings / spouses aboard the Titanic parch # of parents / children aboard the Titanic ticket Ticket number fare Passenger fare cabin Cabin number embarked Port of Embarkation C = Cherbourg, Q = Queenstown, S = Southampton

Variable Notes

pclass: A proxy for socio-economic status (SES)

+ 1st = Upper
+ 2nd = Middle
+ 3rd = Lower

age: Age is fractional if less than 1. If the age is estimated, is it in the form of xx.5

sibsp: The dataset defines family relations in this way...

+ Sibling = brother, sister, stepbrother, stepsister
+ Spouse = husband, wife (mistresses and fiancés were ignored)

parch: The dataset defines family relations in this way...

+ Parent = mother, father
+ Child = daughter, son, stepdaughter, stepson
Some children travelled only with a nanny, therefore parch=0 for them.


###Algorithm
First off, load libraries to use: 

* *dplyr to* data manipulation and commands as group_by and %>% operator
* *ggplot2* and ggthemes for plotting

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggthemes)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
