---
title: "RandomForest"
author: "Diego Polo"
date: "4 de enero de 2018"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TITANIC SURVIVAL FORECASTING USING RANDOMFOREST ALGORITHM

This is my first kernel and I decided to use RamdonForest algorithm as many of features of datasets are factors and I think RandomForest clasificator would work fine

###Overview

The data has been split into two groups:

* training set (train.csv)
* test set (test.csv)

The training set should be used to build your machine learning models. For the training set, we provide the outcome (also known as the “ground truth”) for each passenger. Your model will be based on “features” like passengers’ gender and class. You can also use feature engineering to create new features.

The test set should be used to see how well your model performs on unseen data. For the test set, we do not provide the ground truth for each passenger. It is your job to predict these outcomes. For each passenger in the test set, use the model you trained to predict whether or not they survived the sinking of the Titanic.

We also include gender_submission.csv, a set of predictions that assume all and only female passengers survive, as an example of what a submission file should look like.

###Data Dictionary
VariableDefinitionKey survival Survival 0 = No, 1 = Yes pclass Ticket class 1 = 1st, 2 = 2nd, 3 = 3rd sex Sex Age Age in years sibsp # of siblings / spouses aboard the Titanic parch # of parents / children aboard the Titanic ticket Ticket number fare Passenger fare cabin Cabin number embarked Port of Embarkation C = Cherbourg, Q = Queenstown, S = Southampton

Variable Notes

pclass: A proxy for socio-economic status (SES)

+ 1st = Upper
+ 2nd = Middle
+ 3rd = Lower

age: Age is fractional if less than 1. If the age is estimated, is it in the form of xx.5

sibsp: The dataset defines family relations in this way...

+ Sibling = brother, sister, stepbrother, stepsister
+ Spouse = husband, wife (mistresses and fiancés were ignored)

parch: The dataset defines family relations in this way...

+ Parent = mother, father
+ Child = daughter, son, stepdaughter, stepson
Some children travelled only with a nanny, therefore parch=0 for them.


###Algorithm
First off, load libraries to use: 

* *dplyr to* data manipulation and commands as group_by and %>% operator
* *ggplot2* and ggthemes for plotting

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggthemes)
library(randomForest)
```

Dataset loading and fusion to operate in both dataset
```{r, echo=TRUE}
train=read.csv('./Data/train.csv',stringsAsFactors = FALSE,header = TRUE)
test=read.csv('./Data/test.csv',stringsAsFactors = FALSE,header = TRUE)
full=bind_rows(train,test)
summary(full)
```

###Embarkment

Let's check Embarkment. It shouldn't provide info, it's independent where you embarked and the sunk of Titanic. But I'll check if there are NA in the dataset. Firstly I summarise number of passenger grouped by city of embarkment and lastly number of survivals by city (only in train as I don't know outcomes for test)

```{r,echo=TRUE}
embarked_total=full%>%group_by(Embarked)%>%summarise(num_embarked=n())
print(embarked_total)
embarked_survived=full[1:891,]%>%filter(Survived==1)%>%group_by(Embarked)%>%summarise(num_survived=n())
print(embarked_survived)
```

First, there are 2 NA. I will change them for 'S' as Southhampton is the where most of the people embarked. And I will factorise row
```{r}
full$Mod_Embarked=ifelse(is.na(full$Embarked),'S',full$Embarked)
full$Mod_Embarked=as.factor(full$Mod_Embarked)
```

Let's make some math, I will calculate percentages of these data
```{r, echo=TRUE}
embarked_total=full%>%group_by(Mod_Embarked)%>%summarise(num_embarked=n())
embarked_survived=full[1:891,]%>%filter(Survived==1)%>%group_by(Mod_Embarked)%>%summarise(num_survived=n())

embarked_total$percentage=embarked_total$num_embarked/sum(embarked_total$num_embarked)*100
embarked_survived$percentage=embarked_survived$num_survived/sum(embarked_survived$num_survived)*100
print(data.frame(City=embarked_total$Mod_Embarked, Total=embarked_total$percentage,
                 Survived=embarked_survived$percentage))
```

Data are very similar, i don't think this is a good feature, but people from Cherbourg has higher rates of survival.

###Fares
I continue with 'Fare'. I suppose that higher wealhty, more chances to survive (more boats on decks or not, it can spoil sightseens). Let's do analysis (if it implied Survived I must take full[1,891,])
```{r, echo=TRUE}
summary(full$Fare)

outcomes=full[1:891,]%>%group_by(Fare,Survived)%>%summarise(num_passengers=n())
print(head(outcomes))
```
Lots of fares!! I can't deal with this feature! I will split it in four categories:

+ luxury >300 pounds
+ 200>high>300 
+ 100>medium>200
+ 50>low>100
+ 0>extralow>50

and barplot the result
```{r}
full$Mod_FareLevels=ifelse(full$Fare>=0 & full$Fare<50,'Extralow',ifelse(full$Fare>=50 & full$Fare<100,'Low',ifelse(full$Fare>=100 & full$Fare<200,'Medium',ifelse(full$Fare>=200 & full$Fare<300,'High','Luxury'))))
full$Mod_FareLevels=as.factor(full$Mod_FareLevels)

outcome=full[1:891,] %>% group_by(Mod_FareLevels,Survived) %>% summarise(num_passenger=n())
ggplot(outcome,aes(x=Mod_FareLevels, y=num_passenger,fill = Survived))+
        geom_bar(stat='identity',position='fill', colour='black')+
        labs(y='Passengers', x='Fare',title='Outcomes')+
        ggthemes::theme_few()
```

As I supposed, most expensive fare passengers survived, maybe they were hosted in upper decks and had more time to get a boat.
There's 1 NA value to predict, let's put the most likely value
```{r}
print(full%>%group_by(Mod_FareLevels)%>%summarise(n()))
full$Mod_FareLevels[which(is.na(full$Mod_FareLevels))]='Extralow'
```

###Tickets
This feature is a mess!! I'll take numbers and whether tickets are compound. At a glance compound tickets looks more poshy and expensive.
```{r}
full$Mod_CompoundTicket=as.factor(ifelse(grepl(pattern=" ",x=full$Ticket ),"Yes","No"))
full$Mod_NumberofTicket=sapply(full$Ticket, function(x) strsplit(x, split=" ")[[1]][2])
full$Mod_NumberofTicket=ifelse(is.na(full$Mod_NumberofTicket),full$Ticket,full$Mod_NumberofTicket)
```

I'll check if compound tickets have something special
```{r, echo=FALSE}
outcome=full[1:891,]%>%group_by(Mod_CompoundTicket,Survived)%>%summarise(num_passenger=n())
ggplot(outcome, aes(x=Mod_CompoundTicket,y=num_passenger,fill=Survived))+
    geom_bar(stat="identity",position="fill")+ggthemes::theme_few()
```

It doesn't give us any conclusions... There are some samples without numbers just letters, therefore I'll force them to a 0
```{r}
full$Mod_NumberofTicket=ifelse(grepl(x=full$Mod_NumberofTicket,pattern="[0-9]"),full$Mod_NumberofTicket,"0")
full$NumeroDeTicket=as.numeric(full$Mod_NumberofTicket)
```
 
###Gender
Let's check gender. "Women and children first" must have a meaning in this context.
```{r,echo=FALSE}
full$Sex=as.factor(full$Sex)
outcome=full[1:891,]%>%group_by(Sex,Survived)%>%summarise(num_survival=n())
ggplot(outcome, aes(x=Sex,y=num_survival,fill=Survived))+
    geom_bar(stat='identity',position='fill', colour='Black')+ggthemes::theme_few()
```

Women have more rate of survivals, as expected

###Class
Let's check if class give us any info, expected: higher class more survivals
```{r,echo=FALSE}
full$Pclass=as.factor(full$Pclass)
outcome=full[1:891,]%>%group_by(Pclass,Survived)%>%summarise(num_survival=n())
ggplot(outcome, aes(x=Pclass, y=num_survival,fill=Survived))+
    geom_bar(stat='identity', position='fill', colour='black')+ggthemes::theme_tufte()
```

###Age
Mean age of survivals
```{r}
print(full[1:891,]%>%group_by(Survived)%>%summarise(mean_age=mean(Age,na.rm=TRUE)))
```
Histogram of age distribution

```{r,echo=FALSE}
agemen=full%>%filter(Sex=='male' & !is.na(Age))
agewomen=full%>%filter(Sex=='female' & !is.na(Age))
datamen=data.frame(Age=agemen$Age,Gender='Man')
datawomen=data.frame(Age=agewomen$Age,Gender='Woman')
data=rbind(datamen,datawomen)
ggplot(data, aes(Age,fill=Gender))+geom_histogram(alpha=0.6,binwidth = 5)+ggthemes::theme_fivethirtyeight()
```

No differences detected. Let's classify age in a range: Kid, Teenager, Adult, Elder
```{r}
full$Mod_RangeAge=ifelse(full$Age<1,'Baby',
              ifelse(full$Age<5,'Kid',
              ifelse(full$Age<18, "Teenager",
              ifelse(full$Age<45,"Adult",
              ifelse(full$Age<90,"Elder",
              is.na())))))
```
Another plot to represent survival per age range
```{r,echo=FALSE}
outcome=full[1:891,]%>%group_by(Mod_RangeAge,Sex,Survived)%>%summarise(num_passenger=n())
ggplot(outcome, aes(x=Mod_RangeAge, y=num_passenger,fill=Survived))+
        geom_bar(stat='identity',position='fill')+
        facet_wrap(~Sex)+ggthemes::theme_economist_white()
```

conclusion: if you are male and you are not a child you die. If you are a woman you have more chance to survive.
Being Age so important, it is paramount to predict ages witn NA values. There are `r sum(is.na(full$Age))` NA values. I'll get more features and I will come back later to solve this issue

###Name
I reduce name to only surname as number of names is huge and full of treatment
```{r}
full$Mod_Surname=sapply(full$Name, function(x) strsplit(x,split=",")[[1]][1])
print(head(unique(as.factor(full$Mod_Surname))))
```
I have reduced samples.

###SibSp and Parch
Other feature is wheather they are family or not, for that if SbSp o Parch are different from zero it implies there are relations
```{r,echo=FALSE}
full$Mod_IsFamily=as.factor(ifelse(full$SibSp>0 | full$Parch>0,"Yes","No"))
outcome=full[1:891,]%>%group_by(Mod_IsFamily,Survived)%>%summarise(num_survival=n())
ggplot(outcome, aes(x=Mod_IsFamily, y=num_survival,fill=Survived))+
    geom_bar(stat='identity', position='fill', colour='black')+
    labs(title = 'IsFamily')
    ggthemes::theme_gdocs()

```
Let's examine if there is any correlation between survivals and sibSp and Parch
```{r}
outcome=full[1:891,]%>%group_by(SibSp,Survived)%>%summarise(num_survival=n())
outcome2=full[1:891,]%>%group_by(Parch,Survived)%>%summarise(num_survival=n())
full$Mod_SibSpParch=full$SibSp+full$Parch
outcome3=full[1:891,]%>%group_by(Mod_SibSpParch,Survived)%>%summarise(num_survival=n())
```
Plotting results to analyse
```{r,echo=FALSE}
ggplot(outcome, aes(x=SibSp, y=num_survival,fill=Survived))+
    geom_bar(stat='identity', position='fill', colour='black')+
    labs(title = 'SibSp')
    ggthemes::theme_gdocs()

ggplot(outcome2, aes(x=Parch, y=num_survival,fill=Survived))+
    geom_bar(stat='identity', position='fill', colour='black')+
    labs(title = 'Parch')
    ggthemes::theme_gdocs()
ggplot(outcome3, aes(x=Mod_SibSpParch, y=num_survival,fill=Survived))+
    geom_bar(stat='identity', position='fill', colour='black')+
    labs(title = 'SibSp+Parch')
    ggthemes::theme_gdocs()
```


###Treatmen
```{r}
#extraction of treatments
data=sapply(full$Name, function(x) strsplit(x, split=", ")[[1]][2])
full$Mod_Treatment=sapply(data, function(x) strsplit(x, split="[.]")[[1]][1])
#number of treatments
print(levels(factor(full$Mod_Treatment)))
```
Let's combine male treatment in Mr and female treatment in Mrs or Miss
```{r}
nombresmasculinos=c("Capt","Col","Don","Dr","Jonkheer","Major","Master","Mr","Rev","Sir")
nombresfemeninos=c("Dona","Mlle","Mme","Mrs","Ms","the Countess")

full$Mod_Treatment=ifelse(grepl(paste(nombresmasculinos,collapse="|"),full$Mod_Treatment),"Mr",
               ifelse(grepl(paste(nombresfemeninos,collapse="|"), full$Mod_Treatment),"Mrs",
               "Miss"))
full$Mod_Treatment=as.factor(full$Mod_Treatment)
```

###Age 2
We need to solve NA values in Age. I'll take class, treatment and mean age to define new ages
```{r}
meanages=full%>%group_by(Pclass,Mod_Treatment)%>%summarise(mean_age=mean(Age,na.rm=TRUE))
print(meanages)
#Last item is NaN and put something by default
meanages$mean_age[9]=mean(full$Age[which(full$Mod_Treatment=="Mrs")],na.rm=TRUE)
print(meanages)

full$Mod_Age=full$Age
head(full$Mod_Age)
full$Mod_Age=ifelse(is.na(full$Mod_Age) & full$Pclass==1 & full$Mod_Treatment=="Miss",meanages$mean_age[1],
            ifelse(is.na(full$Mod_Age) & full$Pclass==1 & full$Mod_Treatment=="Mr",meanages$mean_age[2],
            ifelse(is.na(full$Mod_Age) &full$Pclass==1 & full$Mod_Treatment=="Mrs",meanages$mean_age[3],
            ifelse(is.na(full$Mod_Age) &full$Pclass==2 & full$Mod_Treatment=="Miss",meanages$mean_age[4],
            ifelse(is.na(full$Mod_Age) &full$Pclass==2 & full$Mod_Treatment=="Mr",meanages$mean_age[5],
            ifelse(is.na(full$Mod_Age) &full$Pclass==2 & full$Mod_Treatment=="Mrs",meanages$mean_age[6],
            ifelse(is.na(full$Mod_Age) &full$Pclass==3 & full$Mod_Treatment=="Miss",meanages$mean_age[7],
            ifelse(is.na(full$Mod_Age) &full$Pclass==3 & full$Mod_Treatment=="Mr",meanages$mean_age[8],
            meanages$mean_age[9]))))))))

```


I re-apply RangeAge
```{r}
full$Mod_RangeAge=ifelse(full$Mod_Age<1,'Baby',
              ifelse(full$Mod_Age<5,'Kid',
              ifelse(full$Mod_Age<18, "Teenager",
              ifelse(full$Mod_Age<45,"Adult",
              ifelse(full$Mod_Age<90,"Elder",
              '')))))
full$Mod_RangeAge=as.factor(full$Mod_RangeAge)
```

###RandomForest
First check that all features are factors
Let's split data set in train and test subsets
```{r}

full$Survived=as.factor(full$Survived)
full$SibSp=as.factor(full$SibSp)
full$Parch=as.factor(full$Parch)

train=full[1:891,]
test=full[892:nrow(full),]
```
Set random seed
```{r}
set.seed(112)
```
Building the model and compute execution time
```{r}

start_time <- Sys.time()

rf_mod <- randomForest(Survived ~ Pclass+Sex+Mod_FareLevels+Mod_CompoundTicket+Mod_RangeAge+Mod_IsFamily+Mod_Treatment, 
  data = train, 
  ntree = 2000, 
  importance = TRUE)
end_time <- Sys.time()
print(paste0("Training excution time:",end_time - start_time))
```
Plotting errors
```{r}
plot(rf_mod, ylim=c(0,1))
legend('topright', colnames(rf_mod$err.rate), col=1:6, fill=1:6)
```
Get importance of features
```{r}
importance=importance(rf_mod)
print(head(importance))
varImportance=data.frame(Variables = row.names(importance), 
                            Importance = round(importance[ ,'MeanDecreaseGini'],2))
print(head(varImportance))

# Create a rank variable based on importance
rankImportance=varImportance %>%mutate(Rank = paste0('#',dense_rank(desc(Importance))))

ggplot(rankImportance, aes(x = reorder(Variables, Importance),y = Importance)) +
  geom_bar(stat='identity', colour = 'black') +
  geom_text(aes(x = Variables, y = 0.5, label = Rank), hjust=0, vjust=0.55, size = 4, colour = 'lavender',    fontface = 'bold') +
  labs(x = 'Variables', title = 'Relative Variable Importance') +
  coord_flip() + 
  theme_gdocs()
```

###Predictions
Predict using the test set
```{r}
prediction <- predict(rf_mod, test, type = 'response')
# Save the solution to a dataframe
solution = data.frame('PassengerId' = test$PassengerId, 'Survived'=as.numeric(levels(prediction))[prediction])
#solution = data.frame(prediction)
write.csv(solution, 'rf_solution.csv', row.names = FALSE)

```



